Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2018-05-21 paper.tex
Add Yonglin's notes/commentary from email May 16
this updates sections IVB and V

We also solve for the flavor evolution of this system,
integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
as described in \cite{zhu2016-mnr_nsns_remnant}.
In Fig.~\ref{fig:survival-scat}
we show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
for the \emph{scat} treatment. (There is no flavor transformation
in the \emph{noscat} treatment. \footnote{TODO (Brett): ubpdate with new plots})
The survival probability is the probability that a neutrino, if measured,
will be found to be in its original flavor state.
The survival probabilities are computed at each point along the trajectory
from the flavor evolution matrix,
$P_{\nu_\alpha \to \nu_\alpha} = |S_{\alpha \alpha}(E)|^2$.
When $P_{\nu_e \to \nu_e}$ decreases,
as can be seen for example in Fig.~\ref{fig:survival-scat} for $s>400$~km,
some of the electron-type neutrinos have oscillated into
mu- or tau-type neutrinos.

For our assumed matter potential along this one test trajectory
we see that a matter-neutrino resonance only occurs in the \emph{scat} case;
though with a slightly higher matter potential it could occur in both cases,
closer to the disk in the \emph{noscat} case than the \emph{scat} case.
For this particular test trajectory we only see a standard and not a symmetric
matter neutrino resonance;
these two classes of matter-neutrino resonance
are described for example in \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
The difference we see in the oscillation physics between the \emph{noscat}
and \emph{scat} cases along this test trajectory highlights the importance of
scattering physics in the problem of neutrino oscillations in
neutron star mergers.

@@

We have also employed the ray tracing code to examine neutrino flavor
oscillations along one toy trajectory exiting the neutrino-dense environment
of the neutron star post-merger configuration (Sec.~\ref{ssec:V_nunu}).
Along that trajectory,
$25^\circ$ from the polar axis, elastic scattering increases the
$\bar{\nu}_e$ relative to the $\nu_e$ self-interaction potential,
introducing a standard matter neutrino
resonance (MNR) where there was none before.
In a similar configuration, but with matter densities lower
by a factor of a few, the effect of elastic scattering would have been to push
the MNR out to larger radii.
This example demonstrates the importance of the physics of elastic scattering
in the phenomenon of neutrino flavor oscillation.
We caution the reader, however, to avoid drawing general conclusions from this
particular example, since a single astrophysical configuration can present
dramatically different oscillation resonances along test trajectories emerging
at different angles \cite{zhu2016-mnr_nsns_remnant},
and since the MNR is extremely sensitive to a host of parameters.

>>>>

We also solve for the flavor evolution of this system,
integrating $S$ according to Eqn.~\ref{eqn:schroedinger},
as described in \cite{zhu2016-mnr_nsns_remnant}.
We show the survival probabilities for $\nu_e$ and $\bar{\nu}_e$
for the \emph{noscat} treatment in Fig.~\ref{fig:survival-noscat},
and for the \emph{scat} treatment in Fig.~\ref{fig:survival-scat}.
The survival probability is the probability that a neutrino, if measured,
will be found to be in its original flavor state.
The survival probabilities are computed at each point along the trajectory
from the flavor evolution matrix,
$P_{\nu_\alpha \to \nu_\alpha} = |S_{\alpha \alpha}(E)|^2$.
When $P_{\nu_e \to \nu_e}$ decreases,
as can be seen for example in Fig.~\ref{fig:survival-scat} for $s>400$~km,
some of the electron-type neutrinos have oscillated into
mu- or tau-type neutrinos.

In Fig.~\ref{fig:survival-noscat}, we report the survival probabilities
from the calculation with elastic scattering turned off.
Using neutrino mixing angle $\theta_{12}$ and the inverted hierarchy
(the normal hierarchy gives qualitatively similar results)
electron neutrinos and antineutrinos start to oscillate around
700~km in the form of a collective neutrino oscillation,
causing both $\nu_e$ and $\bar{\nu}_e$ to convert to heavy lepton neutrinos
and antineutrinos respectively.
A similar effect was seen in \cite{fren2017-flavor_bns,tian2017-flavor_bns}.
In Fig.~\ref{fig:survival-scat}, we report the survival probabilities
of an otherwise identical calculation, but with elastic scattering turned on.
In this case, electron neutrinos and antineutrinos start to oscillate around
400~km in the form of a standard matter neutrino resonance:
at first both $\nu_e$ and $\bar{\nu}_e$ convert to heavy lepton neutrinos,
but as the transformation progresses, the $\bar{\nu}_e$ partially
return to their original flavor.

Note that during the latter part of the matter neutrino resonance
depicted in Fig.~\ref{fig:survival-scat},
where the self-interaction potential approaches the vacuum scale,
the survival probabilities show some small-scale oscillations
different than the standard matter neutrino resonance
introduced in \cite{malk2015-mnr_2}.
This occurs because the self-interaction potential and matter potential
fall close to the vacuum potential scale.

The collective neutrino oscillation occuring in the \emph{noscat} case
(Fig.~\ref{fig:survival-noscat})
not only produces a very different flavor mixture,
the transformation also starts further from the remnant,
and it extends much further before it completes.
For these potentials along this test trajectory,
we see that a matter-neutrino resonance only occurs in the \emph{scat} case;
though with a slightly higher matter potential it could occur in both cases,
closer to the disk in the \emph{noscat} case than the \emph{scat} case.
For this particular test trajectory we only see a standard and not a symmetric
matter neutrino resonance \cite{malk2016-mnr_3,vaan2016-uncovering_mnr}.
As we can see from these calculations, the final outcome of the flavor
transformation is quite different in the two scenarios.

@@

We have also employed the ray tracing code to examine neutrino flavor
oscillations along one toy trajectory exiting the neutrino-dense environment
of the neutron star post-merger configuration (Sec.~\ref{ssec:V_nunu}).
The trajectory starts from 300~km, and moves out radially at
$25^\circ$ from the polar axis.
Along that trajectory, elastic scattering has the effect of increasing
the ratio of $\bar{\nu}_e$ relative to $\nu_e$.
This creates a negative self-interaction potential
which introduces a complete standard matter neutrino resonance
(see Fig.~\ref{fig:survival-scat}).
At about 400~km from the merger core,
both electron neutrinos and electron antineutrinos begin to transform.
At about 1200~km
electron neutrinos have almost completely converted to mu or tau neutrinos,
while the electron antineutrinos have returned back to their original flavor.
In an otherwise identical calculation, ignoring elastic scattering
causes the flavor transformation to be very different (see Fig.~\ref{fig:survival-noscat}).

This example demonstrates the importance of the physics of elastic scattering
in the phenomenon of neutrino flavor oscillation.
However, we avoid drawing general conclusions from this particular example,
since a single astrophysical configuration can present dramatically different
oscillation resonances along test trajectories emerging at different angles
\cite{zhu2016-mnr_nsns_remnant},
and since the matter neutrino resonance is extremely sensitive to a host
of parameters.
